Biodiversity improves tropical forest resilience

### 2018-11-30 draft.Rmd
Draft 0.1.9008 Final 1 Intro, M&M and 1/2 Res

---
title: Biodiversity improves tropical forest resilience, shifting from complementarity to selection effect
author: "Sylvain Schmitt, Isabelle Maréchaux, Jerome Chave, Fabian Fischer, Camille Piponiot, Stéphane Traissac & Bruno Hérault"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::word_document2: default
  github_document: default
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
    keep_tex: true
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

@@

```{r Ir, cache=TRUE, echo=FALSE}
Ir <- Deq %>% 
  group_by(sim) %>% 
  arrange(time) %>% 
  mutate_at(vars(-time, -sim), cumsum) %>% 
  filter(time == 600)
```

@@

Natural disturbances caused by tree death play a major role in tropical forest functioning through gap-phase regeneration [@Veblen1992]. Climate change and direct human activities are currently deeply modifying forest disturbance regimes [@Davidson2012; @Lewis2015]. If deforestation has been rightly the focus of worldwide attention, forest degradation has been less studied and quantified through tropics [@Putz2010; @Simula2009]. However, degraded forests cover areas over ten times larger than deforestation [@Herold2011], a majority of which is being logged for timber production [@Blaser2011]. Forest degradation can trigger severe biodiversity loss [@Burivalova2014] and is estimated to account for 20.7 billion of tons of carbon dioxide emissions per year [@Pearson2017], representing 20% of greenhouse gas emissions in the Brazilian Amazon [@Asner2005].  In addition to the direct impact of human on forests, a more insidious disturbance is climate change [@Herault2016]. Understanding the drivers of tropical forest resilience to disturbance is thus critical. Here we explored the role of biodiversity in tropical forests answer to disturbances through a simulation approach.

Resilience concept has met numerous definitions, but resilience in ecology can encompass both ecosystem ability to adsorb and recover from disturbance, resistance and recovery respectively [@Gunderson2010]. Both exogenous (e.g. climate, disturbance features) and endogenous (e.g. forest structure and composition) factors can drive ecosystem recovery after disturbance. Among the former, logging intensity has been reported to strongly control and increase time-to-recovery of post-logging tree carbon, subsequent animal biodiversity loss, and change in tree species composition [@Burivalova2014; @DeAvila2015; @Rutishauser2015]. Among endogenous drivers, forest pre-disturbance composition has been identified as a key driver of forest recovery [@Herault2018], as different species respond differently to perturbations depending on their ecological strategies and functional properties [@herault_functional_2011]. Beyond species identity, tree community diversity may also foster forest resilience to perturbations [@DeAvila2018; @Cadotte2011; @Ives2007; @Loreau2013]. Finally, climate change effect on tropical areas results in a significant risk of precipitation decrease [@Davidson2012], whereas intense drought events has been shown to result in huge mortality rates and biomass loss [@Phillips2009].

Two types of mechanisms can underlie an effect of biodiversity on ecosystem properties: complementarity and selection effects [@Loreau2001]. Complementarity effect results from resources partitioning and niche differentiation, and facilitation, leading to a more efficient use of available resources. Selection effect results from a higher probability of more diverse community to include particularly efficient species regarding ecosystem process or stability. Moreover the sampling effect additionally implies redundancy and a lower risk to lose important functional traits, further improving resilience of species rich-forest [@Hooper2005a; @Tilman2001].

The effect of biodiversity on ecosystem functioning has been intensely studied over the last decades, through theoretical work [@Loreau2010], experiments [@Cardinale2007; @Cardinale2009; @Niklaus2017a; @Huang2018], or observations along natural gradient of biodiversity [@Chisholm2013; @Liang2016; @Paquette2011; @Poorter2017]. Whereas several studies advocate only for the importance of few dominant species through selection effect to regulate ecosystem functioning in an old tree diversity experiment [@Bauters2015], in wet or dry natural forests [@Finegan2015; @Prado-Junior2016], or in human-modified forests [@DeAvila2018; @Lohbeck2016]. Numerous authors highlighted diversity role through complementarity, alone or in addition to selection. @Morin2014a identified by simulation an increases of ecosystem stability with species richness. @Fargione2007 observed a shift from sampling to complementarity effect, when @Huang2018 observed an increased complementarity with a null selection effect in forest establishments experiments. And @Cardinale2007; @Poorter2017; @VanderSande2017a reviewed grasslands and forests experiments and study sites and highlighted the role of both complementarity and selection for ecosystem biomass stock and dynamic. However, experiments have mostly been conducted on grasslands [@Hector2005], while studies relying on forest stand inventories often lack repeatability [@Poorter2017; @Schnitzer2016], rarely include disturbed forests [@Sist2015], and monitoring temporal coverage remains too low to assess long-term effect as trees are long-lived organisms [@Herault2018]. Forest simulators have thus proven useful to investigate the effect of biodiversity on forest ecosystem functioning through virtual biodiversity experiments [@Morin2011; @Morin2014a; @Sakschewski2016] However, the problem of biodiversity representation beyond a few plant functional types in models of tropical forests have long been an impediment to apply this approach to this species-rich ecosystem [@Kazmierczak2014; @Kohler1998; @Purves2008].

In the present study, we take advantage of recent developments made in the forest simulator TROLL [@Chave1999], an individual-based and spatially-explicit model of tropical forest, that allow to deal with high diversity levels through a parameterization using plant functional traits [@Marechaux2017]. Consequently, TROLL offers the opportunity to study biodiversity-ecosystem functioning *in silico* in tropical forest communities.

We used a simulation approach to assess the effect of species richness and functional composition on long-term forest recovery from disturbances of different levels. A disturbance can be characterized by a loss of individuals, targeted or not (i.e. species, diameter...), at a specific time, cyclic or not. To have an overview of disturbance effect without confounding effect of disturbance specificity, we used untargeted disturbance happening only once. We manipulated species number and identity across a large number of simulations, and measured the biodiversity net effect on simulated forest recovery after disturbance. In order to further investigate the mechanisms underlying such potential effect of biodiversity, we partitioned it into complementarity and selection effects. Based on the general hypothesis of a positive relationships between biodiversity and productivity [@Liang2016], we hypothesized that more diverse forest will recover quicker to a disturbance event due to an increased productivity [but see @VanderSande2018].

@@

TROLL is an individual-based and spatially explicit forest dynamic model in which the life cycle of individual trees (recruitment, growth, seed production and death) $>1cm$ diameter is simulated. Trees grow in a forest light environment explicitly computed within a $1m^3$ resolution voxel grid. Seedlings below the $1cm$ size class are not modeled explicitly, but are part of a seed/seedlings pool. A maximum of one tree can establish in each $1*1m$ pixel. Each tree is defined by its age, diameter at breast height (dbh), height (h), crown radius (CR), crown depth (CD) and leaf area (LA). Tree geometry is calculated with allometric relationships. Additionally, each tree is flagged with a species label inherited from the parent tree through the seedling recruitment. To each species label is associated a number of species-specific traits (Table \@ref(tab:traits)) from which processes are simulated using up-to-date established relationships drawn from the literature. The model thus represents tree biodiversity in a realistic way, where data requirement and model outputs parallel current trait collection efforts and field inventories, respectively.

Carbon assimilation is computed over half-hourly period of a representative day per month using the Farquhar, von Caemmerer, and Berry model of photosynthesis [@Farquhar1980].  Photosynthetic capacities, as well as leaf respiration rate, are computed from species-specific leaf concentration of nitrogen (N) and phosphorous (P), and leaf mass per area [LMA; @Atkin2015; @Domingues2010]. The net amount of assimilated carbon available for growth is then partitioned among the different plant organs using empirical factors, and converted into increments of stem volume and leaf area using species-specific wood-density and LMA respectively. Leaf demography is simulated using species-specific leaf lifespan derived from LMA. Following an underestimation of leaf lifespan for low LMA species, we designed our own leaf lifespan allometry, specific to tropical tree species, to be used in subsequent analysis (see supplementary materials S1). Natural tree deaths are stochastic events whose background rate is negatively correlated with wood density, as observed pan-tropically [@kraft_functional_2010; @poorter_are_2008; @Wright2010]. Additionally, tree can die due to carbon starvation or treefalls. Below-ground processes, herbaceous plants, epiphytes and lianas are not simulated in this version of TROLL. The source code is written in C++ and available at https://github.com/TROLL-code/TROLL. A further detailed description of the model is available in @Marechaux2017.

@@

A new disturbance module was developed for this study. In order to have a general overview of disturbance effect on forest recovery, we used untargeted disturbance happening only once to be free of confounding factors due to disturbance type specificity. At a given iteration $disturb_{iter}$, individual trees are randomly picked following a uniform law and without any individuals nor species targeting. Selected individuals are then removed without triggering a treefall to avoid any side effect. The operation is repeated until the disturbance result in a defined removed basal area ($disturb_{intensity}$).

@@

In order to assess the role of biodiversity on ecosystem recovery from disturbance, we  simulated forest trajectories for different levels of disturbance intensity and tree community compositions. For each simulation, we first simulated forest regeneration from bare soil for 600 years (pre-disturbance step), then simulated the disturbance described as above ($disturb_{iter}$), and finally let the simulated forest community freely recover from the disturbance during 600 years (post-disturbance step). Disturbance intensity was quantified by the percentage of basal area loss, here fixed at 0% (control), 25%, 50% and 75%. For each level of disturbance, we performed simulations differing in the original number of species ($SR=15,25,125$) and composition. We proceeded in two steps: (i) for each level of species richness, we created 1000 virtual communities by randomly picking desired number of species among the regional species pool; (ii) among those 1000 communities, we selected 20 communities for simulations with a large range of convex hull volume $CVH$ [@Cornwell2006a] but with a community weighted mean close to the regional species pool community weighted mean in order to observe change in functional diversity and not in functional composition. This led to 240 simulations varying in community composition and disturbance intensity ($3~SR*20~CHV*4~ levels~of~disturbance$). We focused on functional richness with convex hull volume $CVH$ and functional mean with community weighted mean $CWM$ to define simulation communities.

@@

@@

Species and functional diversities were assessed for each simulation at the end of the pre-disturbance step with species richness ($SR$) and @villeger_dynamique_2008 indices ($FDiv$, the volume of the functional space occupied by the community, and $FEve$, the regularity of the distribution of abundance in this volume) using the `FD` R package [@laliberte_distance-based_2010-1].

@@

We described the simulated tropical forests from changes in values of forest structure (aboveground biomass ($AGB$ in $ton~C.ha^{-1}$), basal area ($BA$ in $m^2.ha^{-1}$), total number of stem ($N$), number of stem above $10~cm$ diameter ($N10$), and number of stem above $30~cm$ diameter ($N30$)) and in values of forest functioning (growth primary productivity ($GPP$ in $MgC.ha^{-1}$), net primary productivity ($NPP$ in $MgC.ha^{-1}$), tree autotrophic respiration in day ($Rday$ in $MgC.ha^{-1}$) and tree autotrophic respiration in night ($Rnight$ in $MgC.ha^{-1}$)).

The recovery of the simulated forest ecosystem, i.e. the ecosystem capacity to recover from a disturbance event, at a time $t$, $R(t)$ regarding a given property $X$ was defined ad follows (Figure \@ref(fig:IrGraphFinal)):

@@

where $X_D(t)$ and $X_C(t)$ are the ecosystem metric values at time $t$ in the disturbed simulation ($disturb_{intensity} = 25\%, 50\%, 75\%$) and in the undisturbed control simulation ($disturb_{intensity} = 0\%$). Note that demographic stochasticity in the model induces a low variability among replicated simulations with the same set-up so that we did not include repetition of virtual communities simulations [@Marechaux2017].

When $R(t)$ approaches $R_{eq}=1$, the disturbed and control systems are considered equivalent, indicating a perfect recovery. We then calculated the euclidean distance to this state of perfect recovery as $d_{eq}(t) = \sqrt{(R_{eq}-R(t))^2}$. Because the intensity of disturbance was the same for all simulations, our analysis did not include any ecosystem resistance, i.e. the ecosystem capacity to absorb a disturbance event. Consequently in this particular case, we can use the cummulated recovery as an index of ecosystem resilience, further defined as $I_R$ (Figure \@ref(fig:IrGraphFinal)):

\begin{equation}
  I_R(X) = \int _{t=0} ^{t=600} \frac{X_D(t)}{X_C(t)}.dt
  (\#eq:Resilience)
\end{equation}

A high value of $I_R$ indicating a slow recovery. Ecosystem euclidean distance to equilibrium and resilience index $I_R$ was calculated in a multi-dimensional space for the two following forest functions : (i) forest structure ($AGB$, $BA$, $N$, $N10$, and $N30$) and (ii) forest functioning ($GPP$, $NPP$, $Rday$, and $Rnight$). The influence of pre-disturbance species richness ($SR$), functional evenness ($FEve$) and functional diversity ($FDiv$) on ecosystem resilience index $I_R$  was finally investigated by linear regression ($I_R \sim SR + FEve + FDiv$).

```{r IrGraphFinal, fig.cap="Definition of recovery function and resilience indice shown on aboveground biomass dynamics of a control and disturbed ecosystem before and after the disturbance event."}
knitr::include_graphics("images/final.png")
```

@@

The net effect of biodiversity ($NE$) on ecosystem property $X$ is defined as the difference between the ecosystem property $X$ of the mixed-species community and the one expected under the null hypothesis that there is no effect of biodiversity, i.e. ecosystem property $X$ for the mixed-species community equals the abundance-weighted sum of that in monoculture [@Loreau2001]. We then split this net biodiversity effect in complementarity and selection effect using @Loreau2001 partitioning:

\begin{equation}
  \begin{array}{c}
    NE = X_O - X_E = CE + SE \\
    CE = N* \overline{\Delta RX} \overline{M}\\
    SE = N*cov(\Delta RX,M)
  \end{array}
  (\#eq:BiodivPart)
\end{equation}

where $N$ represents the total number of species, and $M$ the vector of monoculture performance. Both metrics depend on the variation of relative ecosystem variable $\Delta RX$:

@@

where $X_{sp}$ is the ecosystem variable value for one species either in mixture $X_{sp}(mixture)$ or in monoculture $X_{sp}(monoculture)$. $P_{sp}$ is the proportion of the species in the mixture represented by species relative abundance. Consequently, $CE$ averages diversity effects of all species presents in the mixture (both negatives and positives). Whereas $SE$ become positive when dominant species outperform themselves in mixture than in monoculture, and negative when less dominant species outperform themselves in mixture than in monoculture [@Tobner2016]. But similarly, to recovery measurement, biodiversity net effect $NE$ is a dynamic equilibrium and vary over time without disturbance. So in order to correctly assess selection and complementarity effect role in tropical forest resilience, we normalized it by ecosystem net effect $NE_C$ from the undisturbed control simulation ($disturb_{intensity} = 0\%$) to measure treatments net effect recovery $R(NE_T)$:

@@

Recovery trajectories of ecosystem variable after disturbance were partitioned between complementarity effect $CE$ and selection effect $SE$. In order to do that, we simulated the 163 species individually representing 652 simulations of monoculture.

@@

## Ecosystem functions

Species richness was not significantly influencing ecosystem resilience index $I_R$ for both forest structure and forest production (Table \@ref(tab:tgIr)). Increased functional diversity reduced  ecosystem resilience index $I_R$ and functional evenness was complementary reducing $I_R$ (Figure \@ref(fig:gIr) and Table \@ref(tab:tgIr)). Finally, low species richness could result in variant $I_R$, but increased species richness resulted in increased functional diversity and consequently lower $I_R$. We found similar results for all disturbance levels (see [Appendix S4: Linear models]).

```{r gIr, echo=FALSE, fig.cap='Ecosystem resilience $I_R$ with taxonomic and functional diversity for different levels of disturbance. Resilience indice $I_R$ was represented against functional diversity [FDiv, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness whereas dot color represents functional evenness [FEve, @villeger_new_2008]. Solid lines and grey areas show fitted relationships with their mean and variance of the form $I_R = FDiv + \\epsilon$.'}
Ir %>% 
  left_join(doe, by = "sim") %>% 
  ungroup() %>% 
  select(disturbance, 
         structure, SR, FEve, FDiv) %>% 
  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), value.name = "Ir") %>% 
  ggplot(aes(FDiv, Ir, shape = SR, col = FEve, group = NA)) +
  geom_smooth(method = 'lm', alpha = 0.3, colour= 'grey') +
  geom_point() +
    facet_wrap(~disturbance, scales = "free", labeller = 'label_both') + 
  xlab('Functional diversity (FDiv)') +
  ylab('Resilience indice Ir') +
  scale_shape_discrete('Species\nrichness\n(SR)') +
  scale_color_continuous('Functional\nevenness\n(FEve)')
```

>>>>

---
title: Biodiversity improves tropical forest resilience
author: "Sylvain Schmitt, Isabelle Maréchaux, Jerome Chave, Fabian Fischer, Camille Piponiot, Stéphane Traissac & Bruno Hérault"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
    keep_tex: true
  github_document: default
  bookdown::word_document2: default
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

@@

```{r Ir, cache=TRUE, echo=FALSE}
Ir <- Deq %>% 
  group_by(sim) %>% 
  arrange(time) %>% 
  mutate_at(vars(-time, -sim), funs(1/cumsum(.))) %>% 
  filter(time == 600)
```

@@

Natural disturbances play a major role in tropical forest functioning through gap-phase regeneration [@Veblen1992]. Climate change and direct human activities are currently deeply modifying forest disturbance regimes [@Davidson2012; @Lewis2015] and while deforestation has been rightly the focus of worldwide attention, the consequences of human-induced disturbance on tropical forests has been less studied [@Putz2010; @Simula2009]. However, degraded forests currently cover areas over ten times larger than deforested ones [@Herold2011], a majority of which is being unsustainably logged for timber production [@Blaser2011]. Forest degradation can trigger severe biodiversity loss [@Burivalova2014] and is estimated to account for 20.7 billion of tons of carbon dioxide emissions per year [@Pearson2017], representing, for instance, 20% of greenhouse gas emissions in the Brazilian Amazon [@Asner2005].  In addition to the direct impact of human on forests, a more insidious disturbance is climate change that modifies regional climates with, for instance, projected increases in the frequency and intensity of convective storms in the Amazon having the potential to increase wind-related tree mortality [@Negron-Juarez2018]. Understanding the drivers of tropical forest resilience to disturbance is thus critical. Here we explored the role of biodiversity in tropical forest recovery after disturbance through a simulation approach.

The concept of ecological resilience has met numerous definitions, but encompass both ecosystem ability to adsorb, the resistance component, and recover from disturbance [@Gunderson2010]. Both exogenous (e.g. climate, disturbance features) and endogenous (e.g. forest structure and composition) factors can drive ecosystem recovery after disturbance. Among the former, logging intensity has been reported to strongly control and increase time-to-recovery of post-logging tree carbon, subsequent animal biodiversity loss, and change in tree species composition [@Burivalova2014; @DeAvila2015; @Rutishauser2015]. Among endogenous drivers, forest pre-disturbance composition has been identified as a key driver of forest recovery [@Herault2018], as different species respond differently to perturbations depending on their ecological strategies and functional properties [@herault_functional_2011]. Beyond species identity, tree community diversity may also foster forest resilience to perturbations [@Cadotte2011; @Ives2007; @Loreau2013]. Finally, disturbance intensity itself has been shown to be a key component of forest resilience [@DeAvila2018].

Two types of mechanisms can underlie the effect of biodiversity on ecosystem properties: complementarity and selection effects [@Loreau2001]. Complementarity effect results from resources partitioning and niche differentiation, and facilitation, leading to a more efficient use of available resources at the community level. Selection effect results from a higher probability of more diverse community to include particularly efficient species regarding ecosystem processes. Moreover the sampling effect additionally implies redundancy and a lower risk to lose important functional traits, further improving resilience of species rich-forest [@Hooper2005a; @Tilman2001].

The effect of biodiversity on ecosystem functioning has been intensely studied over the last decades, through theoretical work [@Loreau2010], experiments [@Cardinale2007; @Cardinale2009; @Niklaus2017a; @Huang2018], or direct observations along natural gradient [@Chisholm2013; @Liang2016; @Paquette2011; @Poorter2017]. Several studies advocate for the importance of few dominant species through selection effect to regulate ecosystem functioning in an old tree diversity experiment [@Bauters2015], in wet or dry natural forests [@Finegan2015; @Prado-Junior2016], or in human-modified forests [@DeAvila2018; @Lohbeck2016]. Numerous authors have, however, highlighted the pivotal role of diversity in ecosystem stability [@Morin2014a] and highlighted the role of both complementarity and selection for ecosystem biomass stock and dynamic [@Cardinale2007; @Poorter2017; @VanderSande2017a]. For instance, @Fargione2007 observed a shift from selection to complementarity effect over time in experimental grasslands, and @Huang2018 observed an increased complementarity over time with no selection effect in forest establishments experiments.

Overall, the role of biodiversity in tropical forest recovery from disturbances and the mechanisms underlying the biodiversity effect on forest functioning along long-term dynamics remain under documented [@VanderSande2017]. Experiments have mostly been conducted on grasslands [@Hector2005], while studies relying on forest stand inventories often lack repeatability [@Poorter2017; @Schnitzer2016], rarely include disturbed forests [@Sist2015], and monitoring temporal coverage remains too low to assess long-term effect as trees are long-lived organisms [@Herault2018]. Forest simulators have thus proven useful to investigate the effect of biodiversity on forest ecosystem functioning through virtual biodiversity experiments [@Morin2011; @Morin2014a; @Sakschewski2016; @Bohn2017]. However, the problem of biodiversity representation beyond a few plant functional types in models of tropical forests have long been an impediment to apply this approach to this species-rich ecosystem [@Kazmierczak2014; @Kohler1998; @Purves2008].

In the present study, we take advantage of recent developments made in the forest simulator TROLL [@Chave1999], an individual-based and spatially-explicit model of tropical forest, that allow to deal with high diversity levels through a parameterization using plant functional traits [@Marechaux2017]. Consequently, TROLL offers the opportunity to study biodiversity-ecosystem functioning  relationships *in silico* in tropical forest communities. More specifically, we used a simulation approach to assess the effect of species and functional diversity on long-term forest recovery from disturbances of different levels. We manipulated species number and identity across a large number of simulations, and measured the biodiversity net effect on simulated forest recovery after disturbance. In order to further investigate the mechanisms underlying such potential effect of biodiversity, we partitioned ecosystem properties into complementarity and selection effects. Based on the general hypothesis of a positive relationships between biodiversity and productivity [@Liang2016], we hypothesized that more diverse forest will recover quicker to a disturbance event due to an increased productivity [but see @VanderSande2018].

@@

TROLL is an individual-based and spatially explicit forest dynamic model in which the life cycle of individual trees (recruitment, growth, seed production and death) $>1cm$ diameter at breast height (dbh) is simulated. Trees grow in a forest light environment computed within a $1m^3$ resolution voxel grid. Seedlings below the $1cm$ size class are not modeled on an individual basis, but are part of a seed/seedlings pool. A maximum of one tree can establish in each $1*1m$ pixel. Each tree is defined by its age, dbh, height (h), crown radius (CR), crown depth (CD) and leaf area (LA). Tree geometry is calculated with allometric relationships. Additionally, each tree is flagged with a species label inherited from the parent tree through the seedling recruitment. To each species label is associated a number of species-specific traits (Table \@ref(tab:traits)) from which ecophysiological processes are simulated using up-to-date established relationships drawn from the literature. The model thus represents tree biodiversity in a realistic way, where data requirement and model outputs parallel current trait collection efforts and field inventories, respectively.

Carbon assimilation is computed over half-hourly period of a representative day per month using the Farquhar, von Caemmerer, and Berry model of photosynthesis [@Farquhar1980].  Photosynthetic capacities, as well as leaf respiration rate, are computed from species-specific leaf concentration of nitrogen (N) and phosphorous (P), and leaf mass per area [LMA; @Atkin2015; @Domingues2010]. The net amount of assimilated carbon available for growth is then partitioned among the different plant organs using empirical factors, and converted into increments of stem volume and leaf area using species-specific wood-density and LMA respectively. Leaf demography is simulated using species-specific leaf lifespan derived from LMA. Following an underestimation of leaf lifespan for low LMA species, we designed our own leaf lifespan allometry, specific to tropical tree species, to be used in subsequent analysis (see supplementary materials S1). Natural tree deaths are stochastic events whose background rate is negatively correlated with wood density, as observed pan-tropically [@kraft_functional_2010; @poorter_are_2008; @Wright2010] an locally [@Aubry-Kientz2013]. Additionally, tree can die due to carbon starvation or treefalls. Below-ground processes, herbaceous plants, epiphytes and lianas are not simulated in this version of TROLL. The source code is written in C++ and available at https://github.com/TROLL-code/TROLL. A further detailed description of the model is available in @Marechaux2017.

@@

A new disturbance module was developed for this study. In order to have a general overview of disturbance effect, we used untargeted disturbance happening only once to be free of confounding factors due to disturbance type specificity. At a given iteration $disturb_{iter}$, individual trees are randomly picked following a uniform law and without any individuals nor species targeting. Selected individuals are then removed without triggering a treefall to avoid any side effect. The operation is repeated until the cummulative basal area of removed trees reaches targeted threshold ($disturb_{intensity}$).

@@

In order to assess the role of biodiversity on ecosystem recovery from disturbance, we  simulated forest trajectories for different levels of disturbance intensity ($disturb_{intensity}$) and tree community compositions. For each simulation, we first simulated forest regeneration from bare soil for 600 years (pre-disturbance step), then simulated the disturbance described as above ($disturb_{iter}$), and finally let the simulated forest community freely recover from the disturbance during 600 years (post-disturbance step). Disturbance intensity, as quantified by the percentage of basal area loss ($disturb_{intensity}$), was fixed at 0% (control), 25%, 50% or 75%. For each level of disturbance, we performed 20 simulations differing in the original number of species (species richness, $SR=15,25,125$) and composition. We proceeded in two steps: (i) for each level of species richness, we created 1000 virtual communities by randomly picking the desired number of species among the regional species pool; (ii) among those 1000 communities, we selected 20 communities for simulations with a large range of convex hull volume $CVH$ [@Cornwell2006a] but with a community mean ($CM$) close to the one of the regional species pool ($CM_{LMA} = 115 \pm 19$, $CM_{wsg} = 0.73 \pm 0.07$, $CM_{dmax} = 0.37 \pm 0.1$, $CM_{hmax} = 40 \pm 11$) in order to get change in functional diversity and not in average functional composition. This led to 240 simulations varying in community composition and disturbance intensity ($3~SR*20~CHV*4~ levels~of~disturbance$). We focused on functional richness with convex hull volume $CVH$ and functional mean with community weighted mean $CWM$ to define simulation communities.

@@

```{r, eval=F}
doe %>% 
  select(starts_with("CWM")) %>% 
  reshape2::melt() %>% 
  group_by(variable) %>% 
  summarise_all(funs(mean, sd, min, max))
```

@@

Species and functional diversities were assessed for each simulation at the end of the pre-disturbance step with species richness ($SR$) and functional diversity indices [$FDiv$, the volume of the functional space occupied by the community, and $FEve$, the regularity of the distribution of abundance in this volume; @villeger_dynamique_2008] using the `FD` R package [@laliberte_distance-based_2010-1].

@@

Simulated tropical forests were characterized in terms of changes in ecosystem properties regarding (i) forest structure (aboveground biomass ($AGB$ in $ton~C.ha^{-1}$), basal area ($BA$ in $m^2.ha^{-1}$), total number of stem ($N$), number of stem above $10~cm$ diameter ($N10$), and number of stem above $30~cm$ diameter ($N30$)) and (ii) forest functioning (growth primary productivity ($GPP$ in $MgC.ha^{-1}$), net primary productivity ($NPP$ in $MgC.ha^{-1}$), tree autotrophic day-time respiration ($Rday$ in $MgC.ha^{-1}$) and tree autotrophic night-time respiration ($Rnight$ in $MgC.ha^{-1}$)).

The recovery of the simulated forest at a time $t$, $R(t)$ regarding a given property $X$ was defined as follows (Figure \@ref(fig:IrGraphFinal)):

@@

where $X_D(t)$ and $X_C(t)$ are the values fo the ecosystem property $X$ at time $t$ in the disturbed simulation ($disturb_{intensity} = 25\%, 50\%, 75\%$) and in the undisturbed control simulation ($disturb_{intensity} = 0\%$) respectively. Note that demographic stochasticity in the model induces a very low variability among replicated simulations with the same set-up so that we we assumed a given simulation is representative of the community properties and dynamics given an initial species pool and disturbance intensity [@Marechaux2017].

When $R(t)$ approaches $R_{eq}=1$, the disturbed and control systems are considered equivalent, indicating a perfect recovery. We then calculated the euclidean distance to this state of perfect recovery as $d_{eq}(t) = \sqrt{(R_{eq}-R(t))^2}$ (Figure \@ref(fig:IrGraphFinal)). Because the intensity of disturbance was set up by the user, our analysis did not include any ecosystem resistance, i.e. the ecosystem capacity to absorb a disturbance event. Consequently in this particular case, we used the inverse of cumulated euclidean distance to the equilibrium as an index of ecosystem resilience, further defined as $I_R$ (Figure \@ref(fig:IrGraphFinal)):

\begin{equation}
  I_R(X) = \frac 1 {\int _{t=disturb_{iter}} ^{t=disturb_{iter}+600} d_{eq}(t).dt}
  (\#eq:Resilience)
\end{equation}

A high value of $I_R$ indicates a fast resilience. Ecosystem euclidean distance to equilibrium and resilience index $I_R$ were calculated in a multi-dimensional space for the two set of ecosystem property values: (i) forest structure ($AGB$, $BA$, $N$, $N10$, and $N30$) and (ii) forest functioning ($GPP$, $NPP$, $Rday$, and $Rnight$).

<!-- $$R[X(t)] = \frac{X_D(t)}{X_C(t)}$$ -->
<!-- $$d_{eq}(t) = \sqrt{(R_{eq}-R(t))^2}$$ -->
<!-- $$I_R(X) = \frac1{\int _{t=disturb_{iter}} ^{t=disturb_{iter}+600} d_{eq}(t).dt}$$ -->

```{r IrGraphFinal, fig.cap="Definition of recovery function R and resilience indice $I_R$ shown on aboveground biomass dynamics for a simulation with $disturb_{intensity}=50%$."}
knitr::include_graphics("images/figure3.png")
```

@@

We tested the influence of pre-disturbance species richness ($SR$), functional evenness ($FEve$) and functional diversity ($FDiv$) on ecosystem resilience index $I_R$ for both forest structure and functioning using one-way ANOVA for species richness ($SR$) and linear regression for functional indices ($I_R \sim FEve + FDiv$). To meet the assumption of normality ecosystem resilience index $I_R$ was log-transformed.

For each ecosystem property $X$, we further investigated the effect of biodiversity and its variation along post-disturbance trajectories. The net effect of biodiversity ($NE$) on ecosystem property $X$ is defined as the difference between the ecosystem property $X$ of the simulated mixed-species community and the one expected under the null hypothesis that there is no effect of biodiversity, i.e. that ecosystem property $X$ for the mixed-species community equals the relative abundance-weighted sum of that in monoculture [@Loreau2001]. In order to do so, we made simulations for each of the 163 species in monoculture and for each level of disturbance leading to 652 simulations of monoculutre.

We then split this net biodiversity effect in complementarity and selection effect using @Loreau2001 partitioning:

\begin{equation}
  \begin{array}{c}
    NE = CE + SE \\
    CE = N* \overline{\Delta RX} \overline{M}\\
    SE = N*cov(\Delta RX,M)
  \end{array}
  (\#eq:BiodivPart)
\end{equation}

where $N$ represents the total number of species, and $M$ the vector of $X$ for the 163 simulations in monoculture and $\Delta RX$:

@@

where $X_{sp}$ is the ecosystem variable value for one species either in mixture $X_{sp}(mixture)$ or in monoculture $X_{sp}(monoculture)$. $P_{sp}$ is the proportion of the species in the mixture represented by species relative abundance. Similarly to recovery measurement, biodiversity net effect $NE$ can vary over time without disturbance. So in order to correctly assess selection and complementarity effect, we normalized it by ecosystem net effect $NE_C$ from the undisturbed control simulation ($disturb_{intensity} = 0\%$) to measure the recovery of the net biodiversity effect $R(NE_T)$:

@@

@@

## Biodiversity effect on forest resilience

Species richness ($SR$) did not significantly influence ecosystem resilience index $I_R$ for forest structure but significantly increased forest functioning resilience (Figure \@ref(fig:gIr) and Table \@ref(tab:tgIr)). Both functional diversity ($FDiv$) and functional evenness ($FEve$) were positively related with $I_R$, i.e. they increased forest resilience (Figure \@ref(fig:gIr) and Table \@ref(tab:tgIr)). Finally, low species richness resulted in variable $I_R$, but increased species richness resulted in increased functional diversity and consequently higher $I_R$ (one-way ANOVA $FDiv \sim SR$, $p-value < 0.001$). We found similar results for all disturbance levels, with increased significance for higher disturbance intensity;

```{r gIr, echo=FALSE, fig.cap='Variation of forest resilience with taxonomic and functional diversity for different levels of disturbance. Resilience indice $I_R$ was represented against functional diversity [FDiv, @villeger_new_2008] for different level of disturbance ($disturb_intensity$} = 25, 50 and 75% of total basal area); dot shapes varies with species richness whereas dot color indicates functional evenness [FEve, @villeger_new_2008]. Solid lines and grey areas show fitted relationships with their mean and variance of the form $I_R = FDiv + \\epsilon$.', fig.width=10, fig.height=15}
Ir %>% 
  left_join(doe, by = "sim") %>% 
  ungroup() %>% 
  select(disturbance, 
         structure, SR, FEve, FDiv) %>% 
  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), value.name = "Ir") %>% 
  ggplot(aes(FDiv, Ir, shape = SR, col = FEve, group = NA)) +
  geom_smooth(method = 'lm', alpha = 0.3, colour= 'grey') +
  geom_point(size = 4) +
  facet_wrap(~disturbance, scales = "free", labeller = 'label_both', 
             nrow = 3) + 
  xlab('Functional diversity (FDiv)') +
  ylab('Resilience indice Ir') +
  scale_shape_discrete('Species\nrichness\n(SR)') +
  scale_color_continuous('Functional\nevenness\n(FEve)') +
  theme(text = element_text(size=20),
        legend.key.size=unit(4,"line"))
```
